---
title: "hw1_643_DieudonneO"
author: "Dieudonne"
date: "June 19, 2016"
output: pdf_document
---




#RECOMMENDER SYSTEM ON MOVIE LENS DATA
#INTRODUCTION
***This is a mini project I wrote for my course Data 643 at CUNY***

*The goal here is to explore recommenderlab,write few functions and predict recommendations to users based on the User-Based -Collaborative-Filtering(UBCF) which through this project appears to be  the most suited*

There are 2 sets of data u.data which is ratings data and u.item data which is movie data

The data are located here http://grouplens.org/datasets/movielens/

```{r,warning=FALSE,message=FALSE,}
#RECOMMENDER SYSTEM ON MOVIE LENS DATA

#(The data are located here http://grouplens.org/datasets/movielens/)
library(recommenderlab)
library(reshape2)

```

#FUNCTION TO GRAB THE DATA 

```{r}

get.Data <- function(){
  
  ##laod ratings data
  ratings <- read.delim("~/Downloads/u.data.txt", header=F)
  colnames(ratings) <- c("userID","movieID","rating", "timestamp")
  
  ## load movies data
  movies <- read.delim("~/Downloads/u.item.txt", sep="|", header=F, stringsAsFactors = FALSE)
  colnames(movies)[colnames(movies)=="V1"] <- "movieID"
  colnames(movies)[colnames(movies)=="V2"] <- "name"
  
  return(list(ratings=ratings, movies=movies))
  
}
```


#FUNCTION FOR  DATA PREPARATION AND PROCESSING

```{r}
Pre.Process = function(ratings, movies)
{ 
  ratings[,2] <- dataList$movies$name[as.numeric(ratings[,2])]
  
  # remove duplicate entries for any user-movie combination
  ratings <- ratings[!duplicated(ratings[,1:2]),]
}
```

#Function to Create movie ratingMatrix from rating Data and movie data 

```{r}
Create.Rating.Matrix <- function(ratings)
{
  # converting the ratingData data frame into rating matrix
  Ratings.Mat <- dcast( ratings, userID ~ movieID, value.var = "rating" , index="userID")
  ratings <- Ratings.Mat[,2:ncol(Ratings.Mat)]
  
  Ratings.Mat.Fin <- as(ratings, "matrix")  ## cast data frame as matrix
  movie.Rating.Mat <- as(Ratings.Mat.Fin, "realRatingMatrix")   ## create the realRatingMatrix
  ### setting up the dimnames ###
  dimnames(movie.Rating.Mat)[[1]] <- row.names(ratings)
  return (movie.Rating.Mat)
}

```


#MODELS

```{r}
evaluateModels <- function(movie.Rating.Mat)
  
{
  ## Find out and analyse available  recommendation algorithm options for realRatingMatrix data
  recommenderRegistry$get_entries(dataType = "realRatingMatrix")
  
  scheme <- evaluationScheme(movie.Rating.Mat, method = "split", train = .9,
                             k = 1, given = 10, goodRating = 4)
  
  algorithms <- list(
    RANDOM = list(name="RANDOM", param=NULL),
    POPULAR = list(name="POPULAR", param=NULL),
    UBCF = list(name="UBCF", param=NULL)
  )
  
  # run algorithms, predict next n movies
  res <- evaluate(scheme, algorithms, n=c(1, 3, 5, 10, 15, 20))
  
  ## select the first results
  
  return (res)
}

```

#VISUALIZATION

```{r}
graphs <- function(res)
{
  # Draw ROC curve
  plot(res, annotate = 1:3, legend="topright")
  
  # See precision / recall
  plot(res, "prec/rec", annotate=3, legend="topright", xlim=c(0,.22))
}

```

#CREATE FUNCTION FOR PREDICTION MODEL

```{r}
create.Model <-function (movie.Rating.Mat,method){
  
  model <- Recommender(movie.Rating.Mat, method = method)
  names(getModel(model))
  getModel(model)$method
  
  getModel(model)$nn
  
  return (model)
}

```

### RATINGS PREDICTIONS USING USER BASED C FILTERING RECOMMENDATIONS


```{r}
rec <- function(movie.Rating.Mat, model, userID, n)
{
  
  ### PREDICT THE TOP N recommendations for given user
  Top.N.List <-predict(model,movie.Rating.Mat[userID],n=n)
  as(Top.N.List,"list")
}

```

# LOAD MOVIE LENS DATA  

```{r}
dataList<- get.Data()
```

# DATA PREPARATION AND PROCESSING

```{r}
ratings<- Pre.Process(dataList$ratings, dataList$movies)
```


# MATRIX FOR MOVIE RATING


```{r}
movie.Rating.Mat<- Create.Rating.Matrix(ratings)

```

# MODELS EVALUTION

```{r}
evalList <- evaluateModels(movie.Rating.Mat)
evalList
```



#The plot for comparing "Random", "Popular", "UBCF" based recommender algorithm is shown:
# plot evaluation result

```{r}
graphs(evalList)

```

## on visualization, looks like UBCF has highest precision.
#The visualisation shows "UBCF" algorithm has highest precision. So I picked "UBCF" to predicts top 10  recommendation of user with userID = 1. 

# get Confusion matrix for "UBCF"


```{r}
getConfusionMatrix(evalList[["UBCF"]])[[1]][,1:4]
```


## LET DO THE RECOMMENDATION BASED ON  "UBCF" 

```{r}

rec_model <- create.Model(movie.Rating.Mat, "UBCF")
userID <- 1
topN <- 5
rec(movie.Rating.Mat, rec_model, userID, topN)
userID<-2
topN<-10
rec(movie.Rating.Mat, rec_model, userID, topN)

```

#Let recommend the top 10 movies for users with ID between 5 and 15 

```{r}

for (userID in 5:15){
  print("We recommend you those movies")
  print(rec(movie.Rating.Mat,rec_model,userID,topN))
}
```




