---
title: "FinalDieudonnefIX1"
author: "Dieudonne"
date: "July 19, 2016"
output: pdf_document
---


#Introduction
**This is the final project of my course Data 643 at CUNY The goal is to explore recommender system in some context as time ,as location ,gender. For this project I am  3 data set all made available by grouplense.**

**For MovieLense and MovieLenseMeta ,'recommenderlab' provides the data** 
**the third data could be dowloaded here**
https://github.com/dieudo/643Summer2016/blob/master/unifiedMLDataMulti.csv

**Initially ,I planed to work and compare packages  avalaible on recommendations systems ,but due to time cnstraint I am going to readjust my goal.**
**This project will be sectioned in 3 part,the first part is comparing and building algorithms around MovieLense data and getting to know the performances associated**.
**The second part is exploring  the users that rated the movies ,can we classified them and learn something related to their age ,their occupations?**
**The third part will be to implement a contextual time value associated to the year**



```{r PART1,warning=FALSE,message=FALSE}
#DATA & libraries
library(plyr)
library(RColorBrewer)
library(grid)
library("recommenderlab")
library(ggplot2)
data_package <- data(package = "recommenderlab")
data_package$results[, "Item"]
data(MovieLense)
str(MovieLense)
str(MovieLenseMeta)
class(MovieLense)
methods(class = class(MovieLense))
data<- read.csv("~/Downloads/unifiedMLDataMulti.csv")
```


#


```{r,warming=FALSE,message=FALSE}
views_per_movie <- colCounts(MovieLense)

views_tbl <- data.frame(
  movie = names(views_per_movie),
  views = views_per_movie
)

views_tbl <- views_tbl[order(views_tbl$views, decreasing = TRUE), ]

head(views_tbl)

ggplot(views_tbl[1:5, ], aes(x = movie, y = views)) +
  geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ggtitle("identity")
ratings_avg <- colMeans(MovieLense)

ratings_avg_relevant <- ratings_avg[views_per_movie > 100]

qplot(ratings_avg_relevant) + stat_bin(binwidth = 0.1) +
  ggtitle(paste("average relevant ratings"))

ratings_movies <- MovieLense[rowCounts(MovieLense) > 50,
                             colCounts(MovieLense) > 100] 
ratings_movies

min_movies <- quantile(rowCounts(ratings_movies), 0.90)
min_users <- quantile(colCounts(ratings_movies), 0.90)

ratings_movies_norm <- normalize(ratings_movies)

# visualize the normalized matrix
image(ratings_movies_norm[rowCounts(ratings_movies_norm) > min_movies,
                          colCounts(ratings_movies_norm) > min_users], main = "with 10 % top")

ratings_movies <- MovieLense[rowCounts(MovieLense) > 50,
                             colCounts(MovieLense) > 100]
ratings_movies


percentage_training <- 0.8
items_to_keep <- 15
rating_threshold <- 3
n_eval <- 1

eval_scheme <- evaluationScheme(data = ratings_movies, method = "split",
                                train = percentage_training, 
                                given = items_to_keep, 
                                goodRating = rating_threshold, 
                                k = n_eval) 
eval_scheme

algorithms_to_evaluate <- list(
  IBCF_cos = list(name = "IBCF", param = list(method ="cosine")),
  IBCF_cor = list(name = "IBCF", param = list(method ="pearson")),
  UBCF_cos = list(name = "UBCF", param = list(method ="cosine")),
  UBCF_cor = list(name = "UBCF", param = list(method ="pearson")),
  random = list(name = "RANDOM", param = NULL)
)

n_recommendations <- c(1, 5, seq(10, 100, 10))

results <- evaluate(eval_scheme, algorithms_to_evaluate, type = "ratings")
plot(results)

sapply(results, class) == "evaluationResults"
lapply(results, avg)

sapply(results, avg)

recom_results <- evaluate(x = eval_scheme, method = algorithms_to_evaluate, n = n_recommendations)

sapply(recom_results, class) == "evaluationResults"
avg_matrices <- lapply(recom_results, avg)

plot(recom_results, annotate = c(1,3,5),  legend = "topleft") 
title("ROC curve")

#UBCF_COS appeared to perform well in this dataset 

```

**UBCF_COS appeared to perform well in this dataset** 


#PART2 EXPLORING THE SECOND DATA


```{r PART2,warning=FALSE,message=FALSE}

library(RColorBrewer)
library(grid)
library(plyr)
library(dplyr)

ggplot(data, aes(x=as.factor(rating),y=age)) +
  geom_boxplot(aes(fill=as.factor(rating))) +
  scale_fill_discrete(name="rating") +
  facet_wrap(~occupation)+xlab("rating")

# age VS rating
ggplot(data, aes(x=as.factor(rating),y=age)) +
  geom_boxplot(aes(fill=as.factor(rating))) +
  scale_fill_discrete(name="rating") 

# rating VS gender
gender.df <- as.data.frame(table(data$gender, data$rating))
ggplot(gender.df, aes(x=Var1, y=Var2)) +
  geom_tile(aes(fill = Freq), colour = "black") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  xlab("gender") +
  ylab("rating")

# rating VS occupation
occupation.df <- as.data.frame(table(data$occupation, data$rating))
ggplot(occupation.df, aes(x=Var1, y=Var2)) +
  geom_tile(aes(fill = Freq), colour = "black") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  xlab("occupation") +
  ylab("rating")+coord_flip()

# rating VS genre
genre.df <- as.data.frame(table(data$genre, data$rating))
ggplot(genre.df, aes(x=Var1, y=Var2)) +
  geom_tile(aes(fill = Freq), colour = "black") +
  scale_fill_gradient(low = "white", high = "steelblue") +
  xlab("genre") +
  ylab("rating")+coord_flip()

# age VS rating VS occupation
ggplot(data, aes(x=as.factor(rating),y=age)) +
  geom_boxplot(aes(fill=as.factor(rating))) +
  scale_fill_discrete(name="rating") +
  facet_wrap(~occupation)

```

#Age component 

**young people rate lower ,rating seems to be positively related to age .**

#Gender

**Men rate more than women and they rates at 4 most**

#Occupation

**Comparing with other occupations, the number of students who rate is the largest; and students rates at 4 most**.

#Type of movie

**Drama, comedy, action have more rates**


#PART3  TIME CONTEXT IN TERM OF THE YEAR RELEASE

```{r PART3,warning=FALSE,message=FALSE}
#summary(MovieLenseMeta)
#str(MovieLense)
df=MovieLenseMeta
#We will only consider users  that views more than 100 movies and who rate more than 50 movies
ratings_movies <- MovieLense[rowCounts(MovieLense) > 50, colCounts(MovieLense) > 100]



set.seed(0)
test <- sample(x = 1:5,
                    size = nrow(ratings_movies),
                    replace = TRUE)
for(i in 1:5) {
  train <- test == i
  Rtrain <- ratings_movies[train, ]
  Rtest <- ratings_movies[!train, ]
}

model<- Recommender(data = Rtrain, method = "UBCF")

predictions <- predict(model, newdata = Rtest, n = 15)

#Choose the biggest year as a benchmark
m <- max(MovieLenseMeta$year, na.rm = TRUE)
#Check the number of years between

n.y <- m- df$year

yrs <- as.numeric(levels(as.factor(n.y)))
wts <- 1 / log(yrs + exp(1))

#year weight
y.w <- 1 / log(n.y + exp(1))
#Exponential decay
plot(df$year,y.w,col="blue",main="by year")
plot(n.y,y.w,col="red",main="by the difference in time")
#Remove na 
y.w[is.na(y.w)] <- 0
weights <- data.frame(title = df$title, wt = y.w, stringsAsFactors = FALSE)

Rdf <- data.frame(user = sort(rep(1:length(predictions@items),predictions@n)), rating = unlist(predictions@ratings), index = unlist(predictions@items))
Rdf$title <- predictions@itemLabels[Rdf$index]
library(dplyr)
Rwt <- inner_join(Rdf, weights, by = "title")

Rwt <- Rwt%>% mutate(wt_rating = rating * wt) %>% group_by(user) %>% arrange(desc(wt_rating)) %>% select(user, title, wt_rating) %>% top_n(5)
head(Rwt, 13)

Rdf2 <- Rdf %>% group_by(user) %>% arrange(desc(rating)) %>% select(user, title, rating) %>% top_n(5, rating)

head(Rdf2, 10)
```

#CONCLUSION AND FUTURE WORK

**From this work ,we can see that the best recommender algorithm depend on the data we have in hand ,the context of the business will determine which algorithm to pick.One must dive deeper to understand the paterns of the data as well to get insights  ,it is also important to explore all types of data that could get involved in the business .**
**It is also important to take into consideration other factors such as time ,location and connections to other intities.**
**For future work ,I would  to extend this work on computing similarity related to the age ,sex,occupation and give recommendations based on those factors.** 



Reference : 
Suresh K. Gorakala, Michele Usuelli
**Time weights**
Building a Recommendation System with R
https://www.packtpub.com/big-data-and-business-intelligence/building-recommendation-system-r

 **unifiedMLDataMulti.csv**
Justin Chu : cleaning the grouplense data and puting into :unifiedMLDataMulti.csv

https://github.com/JustinChu?tab=repositories
