---
title: "project4.643"
author: "Dieudonne"
date: "July 14, 2016"
output: pdf_document
---

#This the fourth mini Project for my course 643 at CUNY
# In general ,businesses have customers database where you can find their past purchase's history but usually there  is no  actual ratings associated to those purchases .How can we build some  kind of ratings based on  the quantities purchased and generate  a recommender system that could be profitable to a company ?Answering this question is the goal in this project .For this particular assignment I use viavi solutions Quaterly Sales dataset .But this can be generalize to many other cases where we can identify customers,items purchased(or service provided) and the amount or quantity purchased 
https://github.com/dieudo/643Summer2016/blob/master/QuarterlySalesProject4.csv
http://www.viavisolutions.com/en-us 

```{r,message=FALSE,warning=FALSE}

library(recommenderlab)
library(reshape2)
library(ggplot2)
# Read training file along with header
library(arules)
library(recosystem)
#library(SlopeOne)
#library(SVDApproximation)
library(knitr)
library(data.table)
library(RColorBrewer)
library(ggplot2)
df<- read.csv("~/Downloads/QuarterlySalesProject4.csv")
library(psych)
#describe(tr)
head(df)
attach(df)
table(Quantity)
str(Quantity)
hist(table(Quantity),col="red")
hist(Quantity,col="blue")
names(df)
str(df)
summary(df)
g<-acast(df, Customer ~ Item)
# Check the class of g
class(g)
 
```

# Matrix convertion

```{r}
R<-as.matrix(g)
 
# Convert R into realRatingMatrix data structure
#   realRatingMatrix is a recommenderlab sparse-matrix like data-structure
r <- as(R, "realRatingMatrix")
 
# I can turn it into data-frame
kable(head(as(r, "data.frame")))
 
```


# The ratings matrix need to be normalized 

```{r}

r_m <- normalize(r)

 
image(r[1:100], main = "Raw Ratings")    
image(r_m[1:100], main = "Normalized Ratings")
 
rb <- binarize(r, minRating=1)
#head(as(rb, "matrix"))
image(rb, main = "binarized  Ratings")  


```

# Modeling and algorithms and similarity measure

```{r}
#UBCF: User-based collaborative filtering
#
model1=Recommender(r[1:nrow(r)],method="UBCF", param=list(normalize = "Z-score",method="Cosine",nn=5, minRating=1))
model2=Recommender(r[1:nrow(r)],method="UBCF", param=list(normalize = "Z-score",method="Jaccard",nn=5, minRating=1))

#IBCF: Item-based collaborative filtering
#
model3=Recommender(r[1:nrow(r)],method="IBCF", param=list(normalize = "Z-score",method="Jaccard",minRating=1))
# POPULAR
#
model4=Recommender(r[1:nrow(r)],method="POPULAR")
 

print(model3)
names(getModel(model3))
getModel(model3)$nn

print(model1)
names(getModel(model1))
getModel(model1)$nn

```


#Predictions and Recommendations to particular customers using differents models

```{r}
# Recommendation to  comcast using model 3
#
Rec.comcast3 <- predict(model3, r["COMCAST",], n=5)

#Top 4  using model4
Rec.comcast4 <- predict(model4, r["COMCAST",], n=10)

Best3comcast <- bestN(Rec.comcast4, n = 3)
Best3comcast
as(Best3comcast, "list")

#Recommendation to JAS 
#
Rec.JAS <- predict(model3, r["JAS",], n=10)
Best5JAS<-bestN(Rec.JAS,n=5)
as(Best5JAS, "list")
recom <- predict(model3, r[1:nrow(r)], type="ratings")
recom
 
```

#Models examination

```{r}

# Convert all your recommendations to list structure
rec_list<-as(recom,"list")
head(summary(rec_list))
# Access service and ratings(type and quantity purchased)with id 1 and first item on the list
rec_list[[1]][1]
# Convert to data frame all recommendations for user 1
Udf<-as.data.frame(rec_list[[1]])
#Display other services to be recommended to the same  customer 
head(attributes(Udf),10)
Udf$id<-row.names(Udf)
head(Udf)

evals <- evaluationScheme(r, method="cross-validation",
        k=4, given=1)
evals
# creation of recommender model based on ubcf
Rec.ubcf <- Recommender(getData(evals, "train"), "UBCF")
# creation of recommender model based on ibcf for comparison
Rec.ibcf <- Recommender(getData(evals, "train"), "IBCF")
# making predictions on the test data set
p.ubcf <- predict(Rec.ubcf, getData(evals, "known"), type="ratings")
# making predictions on the test data set
p.ibcf <- predict(Rec.ibcf, getData(evals, "known"), type="ratings")
# obtaining the error metrics for both approaches and comparing them
error.ubcf<-calcPredictionAccuracy(p.ubcf, getData(evals, "unknown"))
error.ibcf<-calcPredictionAccuracy(p.ibcf, getData(evals, "unknown"))
error <- rbind(error.ubcf,error.ibcf)
rownames(error) <- c("UBCF","IBCF")
error
scheme <- evaluationScheme(r[1:500], method="cross", k=4, given=1,goodRating=5)
results <- evaluate(scheme, method="POPULAR", n=c(1,3,5,10,15,20))
getConfusionMatrix(results)[[1]]
avg(results)
plot(results, annotate=TRUE)

head(as(recom, "matrix")[5,3])   # Rating for user 5 for item at index 3
head(as.integer(as(recom, "matrix")[5,3]))# Just get the integer value
head(as.integer(round(as(recom, "matrix")[9,8]))) # Just get the correct integer value
head(as.integer(round(as(recom, "matrix")[8,7]))) 
 
# Convert recommendations to list structure
rec_list<-as(recom,"list")
head(summary(rec_list))
```

